### 2016-10-03 FrontBackmatter/Acknowledgments.tex
Luca's comments and corrections are in for the main body of the thesis
Going through typos and typography

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

\epigraph{
Just a perfect day

You made me forget myself

I thought I was

Someone else, someone good}{
--- \textsc{Lou Reed}}

\endgroup

>>>>

\chapter*{Acknowledgments}

It has not been one day, it's been four years.\footnotetext{
Just a perfect day \\
You made me forget myself \\
I thought I was \\
Someone else, someone good \\
\hspace*{15pt}
--- \textsc{Lou Reed}}
And it's not always been perfect.
It is however true that the people I have met and interacted with during
this adventure have made me feel \emph{someone good}.
I don't feel it's an exaggeration to say that if I have made it this far
is also because of \emph{you}.
Thus, these few, poorly written lines are
a \emph{dedication} of this dissertation to you
and an \emph{acknowledgment}
of the patience and love you have shown towards me.
Thanks for not giving up on me.

First and foremost, \textitalian{ai miei genitori,
Adriano e Vilma. Ai miei fratelli, Francesca e Simone. A mia zia, Marina
ed ai miei nonni, Cherubino e Irma, Maria e Piero. Questo è dedicato a
voi, per tutto quello che mi avete insegnato, per il sostegno continuo e
la pazienza infinita che dimostrate ogni giorno.}

in no particular order:\footnotetext{
And all of my friends were there, \\
and no one cared \\
\hspace*{15pt}
--- \textsc{The Kinks}}
\begin{itemize}
  \item To \emph{Radovan Bast}, \emph{il miglior fabbro},
  \item To \emph{Davide Michetti}, the inventor of nicknames,
  \item To \emph{The Little Lebowski Urban Achievers}, Erlend, Ana and
    Torgeir,
  \item To the \emph{Schrodinger killed my cat!}
\end{itemize}

### 2016-10-03 FrontBackmatter/Introduction.tex
Going through typos and typography

\begin{epigraphs}
\qitem{
[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
}{
--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}}
\end{epigraphs}

\section*{What Are the Electrons Really Doing in Molecules?}

@@

\arrowthought{It is easy to manipulate chemicals in a virtual laboratory.} Nowadays
quantum chemical methods often complement traditional experimental
approaches. They have become invaluable tools in the modern development
of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye} as
witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998} and
2013.\autocite{Nobel2013}
The concept of \emph{model chemistries} is at the heart of these
successes. Introduced by \citeauthor{Pople1999-gt}, theoretical model
chemistries are specific combinations of approximations in the basis set
and molecular electronic structure methodology.\autocite{Pople1999-gt}
Model chemistries are \emph{systematically improvable} so that it is
possible to achieve the heaven of chemical accuracy by relaxing
approximations in the model, albeit at an increased computational cost.
Figure \ref{fig:RQC-axis} presents the usual depiction of this concept
as a set of orthogonal axes, where chemical accuracy can be achieved by
moving away from the origin.
Relativity can be considered as the third axis of quantum
chemistry: accuracy can be improved on the Hamiltonian, method
and basis axes.\autocite{Saue2011-qg}

The description of the ideas and methods at the basis of quantum
chemistry will be the subject of Chapter \ref{ch:QM}.
I will put emphasis on the methods that have been relevant in the work
presented in this dissertation. Section \ref{sec:mqm} discusses some of the
ideas of quantum many-body theory relevant to quantum chemistry. Section
\ref{sec:mean-field} introduces the workhorses of quantum chemistry
\acrlong{HF} and \acrlong{KS} \acrlong{DFT}.
The \acrlong{MBPT} and \acrlong{CC} rungs on the methodology axis will
be discussed in Sections \ref{sec:mbpt} and \ref{sec:coupled-cluster},
respectively.

@@

\newthought{Before introducing our model of choice}, let us consider how an
environment might affect molecular observables of interest.
Environment effects are usually classified as:
\begin{description}[leftmargin=2mm, font=\normalfont]
\item[Direct.]
  These effects stem straightforwardly from the modification underwent by
  the solute electronic density when interacting with the environment.
\item[Indirect.]
  It is common for solutes to exhibit different minimum-energy
  conformations in different environments. These effects are commonly
  labelled as indirect.
\item[Local field.]
  Light-matter interactions are also affected by the environment. Local
  modifications of externally applied fields subtly influence molecular
  responses.\autocite{Cammi1998-jp, Pipolo2014-sd}
\item[Dynamic.]
  The presence of the environment radically influences excited states,
  since relaxation processes in the medium become
  important.\autocite{Aguilar1993-pu, Cammi1995-ah}
\item[Specific.] This catch-all category includes all effects
  stemming from the peculiar solute-solvent pair interactions that
  cannot be fully described under any of the previous labels.
  In general, modelling such effects demands an atomistic level of
  detail.
\end{description}

@@

\section*{The Road to Reality or Molecular Response Properties}

\begarrowthought{The scope of quantum chemistry is to understand and predict molecular
properties}, but so far we have made no mention of properties other than
the energy.
The experimentalists' view of molecular systems is built mainly around
the use of spectroscopic techniques that explore the interaction of
light and matter. When a system is exposed to an external perturbing
electromagnetic field, it will respond with a detectable change in its
properties.\autocite{Pedersen2012-il, Jaszunski2012-wy}
Characterizing, explaining and predicting a large number of measurable
properties requires a synergistic experimental and theoretical approach.
\emph{Response theory} is the missing link between theory and
experiment, making quantum chemistry a full-fledged virtual laboratory.
Response theory allows the description and computation of
perturbation-induced changes in observable molecular properties.
Electric and magnetic properties, excitation energies and transition
moments can easily be calculated in the framework of response theory.
\emph{Response functions} are the central concept in response theory.
These are built solely by means of \emph{unperturbed} molecular states and
energies: no explicit modelling of excited states is needed.

@@

\begarrowthought{The concepts of systematic improvability and theoretical model
chemistries} are the foundations for the successful practice of quantum
chemistry.
A balanced description of electron correlation is often necessary
to achieve accurate enough results in the computation of molecular
properties.\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye}
The cost-effective treatment of electron correlation is a challenging
problem and a very active line of research in the field.
\Acrlong*{DFT} approaches are cheap and widespread, but their general
accuracy is hard to assess. \Acrlong*{MBPT} and \acrlong{CC} approaches
are more robust in this respect, albeit at an increased computational
cost.

In this dissertation, we are interested in the inclusion of
environment effects in \emph{ab initio} models of interesting chemical
systems. As already noted, this is a challenging problem, the more so
when including electron correlation is necessary to obtain better
accuracy.
We will describe our approach to this problem in Chapter
\ref{ch:solvation-correlation}. Treating solvation and electron
correlation has been a recurring subject of research in the literature
since the inception of continuum models. Once again, we will leverage
the variational formulation of classical polarizable models presented
in Chapter \ref{ch:CSM}.

>>>>

\section*{What Are the Electrons Really Doing in Molecules?}\footnotetext{
[...] \textnorsk{og fordi jeg alltid har hatt en dragning mot det skjulte og hemmelige.}
\\
\hspace*{15pt}
--- \textnorsk{\textsc{K. O. Knausgård}, \textit{Om Høsten}}
}

@@

\arrowthought{It is easy to manipulate chemicals in a virtual laboratory.} Nowadays
quantum chemical methods often complement traditional experimental
approaches. They have become invaluable tools in the modern development
of chemistry,\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye} as
witnessed by the Nobel prizes awarded in 1998\autocite{Nobel1998} and
2013.\autocite{Nobel2013}
The concept of \emph{model chemistries} is at the heart of these
successes. Introduced by \citeauthor{Pople1999-gt}, theoretical model
chemistries are specific combinations of approximations in the basis set
and molecular electronic structure methodology.\autocite{Pople1999-gt}
Model chemistries are \emph{systematically improvable} so that it is
possible to achieve the heaven of chemical
accuracy\footnote{A theoretical result is said to be \emph{chemically
accurate} if it is within \SI{1}{\kilo\calorie\per\mol} of the corresponding
experimental measurement.} by relaxing
approximations in the model, albeit at an increased computational cost.
Figure \ref{fig:RQC-axis} presents the usual depiction of this concept
as a set of orthogonal axes, where chemical accuracy can be achieved by
moving away from the origin.
Relativity can be considered as the third axis of quantum
chemistry: accuracy can be improved on the Hamiltonian, method
and basis axes.\autocite{Saue2011-qg}

The description of the ideas and methods at the basis of quantum
chemistry will be the subject of Chapter \ref{ch:QM}.
I will put emphasis on the methods that have been relevant in the work
presented in this dissertation.

@@

\newthought{Before introducing our model of choice}, let us consider how an
environment might affect molecular observables of interest.
Environment effects are usually classified as:
\begin{description}[leftmargin=2mm, font=\itshape]
\item[Direct.]
  These effects stem straightforwardly from the modification underwent by
  the solute electronic density when interacting with the environment.
\item[Indirect.]
  It is common for solutes to exhibit different minimum-energy
  conformations in different environments. These effects are commonly
  labelled as indirect.
\item[Local field.]
  Light-matter interactions are also affected by the environment. Local
  modifications of externally applied fields subtly influence molecular
  responses.\autocite{Cammi1998-jp, Pipolo2014-sd}
\item[Dynamic.]
  The presence of the environment radically influences excited states,
  since relaxation processes in the medium become
  important.\autocite{Aguilar1993-pu, Cammi1995-ah}
\item[Specific.] This catch-all category includes all effects
  stemming from the peculiar solute-solvent pair interactions that
  cannot be fully described under any of the previous labels.
  In general, modelling such effects demands an atomistic level of
  detail.
\end{description}

@@

\section*{A Road to Reality or Molecular Response Properties}

The experimentalists' view of molecular systems is built mainly around
the use of spectroscopic techniques that explore the interaction of
light and matter. When a system is exposed to an external perturbing
electromagnetic field, it will respond with a detectable change in its
properties.\autocite{Pedersen2012-il, Jaszunski2012-wy}
Characterizing, explaining and predicting a large number of measurable
properties requires a synergistic experimental and theoretical approach.
\emph{Response theory} is the missing link between theory and
these vast classes of
experiments, making quantum chemistry a full-fledged virtual laboratory.
Response theory allows the description and computation of
perturbation-induced changes in observable molecular properties.
Electric and magnetic properties, excitation energies and transition
moments can easily be calculated in the framework of response theory.
\emph{Response functions} are the central concept in response theory.
These are built solely by means of \emph{unperturbed} molecular states and
energies: no explicit modelling of excited states is needed.

@@

\begarrowthought{The concepts of systematic improvability and theoretical model
chemistries} are the foundations for the successful practice of quantum
chemistry.
A balanced description of electron correlation is often necessary
to achieve results that are accurate enough for meaningful comparisons
with experiments.\autocite{Lee1995-pw, Helgaker2004-oz, Tajti2004-ye}
The cost-effective treatment of electron correlation is a challenging
problem and a very active line of research in the field.
\Acrlong*{DFT} approaches are cheap and widespread, but their general
accuracy is hard to assess. \Acrlong*{MBPT} and \acrlong{CC} approaches
are more robust in this respect, albeit at an increased computational
cost.

In this dissertation, we are interested in the inclusion of
environment effects in \emph{ab initio} models of interesting chemical
systems. As already noted, this is a difficult problem, the more so
when including electron correlation is necessary to obtain better
accuracy.
We will describe our approach to this problem in Chapter
\ref{ch:solvation-correlation}. Treating solvation and electron
correlation has been a recurring subject of research in the literature
since the inception of continuum models. Once again, we will leverage
the variational formulation of classical polarizable models presented
in Chapter \ref{ch:CSM}.
